---
title: "Analyses with ESS 9 data to reproduce N/I"
author: "Armin Schäfer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

In this file, I use round 9 ESS data to reproduce some of the analyses of Norris and Inglehart (2019) "Cultural Backlash" and also provide my own additional analyses.

Packages and functions
<<<<<<< HEAD
```{r, message = FALSE}
=======
```{r}
>>>>>>> bf4084247c05632565b8e98746a0fd6754471bc0
# Loading R-libraries.

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, sjlabelled, data.table, psych,
               here, patchwork, modelsummary)

tabl <- function(...) table(..., useNA='ifany')

```


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

Load the data
```{r}

ess <- readRDS(here("processed-data", "arb_ess9.RDS"))

```

Political trust, perceived responsiveness and RWP-Vote
```{r}


p1 <- ess %>%
  group_by(class16_fac) %>%
  filter(!is.na(class16_fac)) %>%
  mutate(trust = 100-trust) %>%
  summarise(pop =  weighted.mean(popauth, weight, na.rm=T),
            trustinst = weighted.mean(trust, weight, na.rm=T))%>%
ggplot(aes(x = trustinst, y = pop))+
  geom_smooth(method="lm", color="grey70", alpha=0.3)+
  geom_point(size=2)+
  ggrepel::geom_text_repel(aes(label=class16_fac), size=3)+
  #scale_y_continuous(labels = scales::percent)+
  labs(x = "\nPolitical trust",
       y = "Populist-authoritarian vote\n",
       shape = NULL)+
  theme_bw()+
  theme(axis.text = element_text(color = "black"),
        axis.title = element_text(face = "bold"))

p2 <- ess %>%
  group_by(class16_fac) %>%
  filter(!is.na(class16_fac)) %>%
  summarise(pop = weighted.mean(popauth, weight, na.rm=T),
            respo = weighted.mean(response, weight, na.rm=T))%>%
  ggplot(aes(x = respo, y = pop))+
  geom_smooth(method="lm", color="grey70", alpha=0.3)+
  geom_point()+
  ggrepel::geom_text_repel(aes(label=class16_fac), size=3)+
  #scale_y_continuous(labels = scales::percent,
  #                  limits = c(0,0.25))+
  labs(x = "\nPerceived responsiveness",
       y = "Populist-authoritarian vote\n",
       shape = NULL)+
  theme_bw()+
  theme(axis.text = element_text(color = "black"),
        axis.title = element_text(face = "bold"),
        axis.text.y = element_blank(),
        axis.title.y = element_blank())

p1 + p2
ggsave(here("figures", "vote_trust_resp.png"),
       device="png", height=6, 
       width=12, units="in", dpi=800)

```


<<<<<<< HEAD
Regression analysis - authoritarian populist parties.

=======
Regression
>>>>>>> bf4084247c05632565b8e98746a0fd6754471bc0
```{r}
library(ggeffects)
library(lme4)

m1 <- lm(popauth~ cohort, data = ess, weights=weight)
m2 <- glm(authpop~ cohort + married + separate + factor(class5) + edu5 + gndr + urban + 
            relig + hincfel + uemp12m + schwartzauth, data = ess, family = "binomial", weights=weight)
m3 <- lm(popauth~ cohort + married + separate + factor(class5) + edu5 + gndr + urban + 
            relig + hincfel + uemp12m + schwartzauth +
            + immig + trust, data = ess, weights=weight)

             
summary(m3)

ggpredict(m2, c("cohort"))

<<<<<<< HEAD
=======


>>>>>>> bf4084247c05632565b8e98746a0fd6754471bc0
ggplot(data = predage, aes(x = x, y = predicted))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, color=NULL),
              alpha=0.4)+
  geom_line(color="black")+
  ylim(0,1)+
  labs(alpha = NULL,
       fill = "Bildung",
       x = "Alter",
       y = "Vorhergesagte Beteiligung \n")+
  theme_minimal()+
  theme(axis.text = element_text(color = "black"))

```

<<<<<<< HEAD
Regression analyses authoritarian populist parties and plotting the coefficients.
=======
>>>>>>> bf4084247c05632565b8e98746a0fd6754471bc0
```{r}

mod <- lm(popauth~ cohort + gender + married + separate + children + 
             class5 + edu5 + ethnic + relig + unemp + lrscale + diseco + 
             schwartz01 + trust01 + immig01 +  factor(cntry), weights=weight, data = ess)


cm <- c('cohortBaby Boomers' = 'Baby Boomers',
        'cohortGen X' = 'Generation X',
        'cohortMillennials' = 'Millennials',
        'schwartz01' = 'Authoritarian values',
        'trust01' = 'Political trust')

p1 <- modelplot(mod, coef_map= cm)+
  geom_vline(xintercept = 0, color="grey70", alpha=0.7)+
  xlab("Regression coefficients")+
  scale_y_discrete(labels = scales::wrap_format(14))+
  theme_bw()+
  theme(axis.text = element_text(color="black"))


mod2 <- lm(popauth~ cohort + gender + married + separate + children + 
             class5 + edu5 + ethnic + relig + unemp + lrscale + diseco + 
             schwartz01 + response01 + immig01 + factor(cntry), weights=weight, data = ess)


cm <- c('cohortBaby Boomers' = 'Baby Boomers',
        'cohortGen X' = 'Generation X',
        'cohortMillennials' = 'Millennials',
        'schwartz01' = 'Authoritarian values',
        'response01' = 'Perceived responsiveness')

p2 <- modelplot(mod2, coef_map= cm)+
  geom_vline(xintercept = 0, color="grey70", alpha=0.7)+
  xlab("Regression coefficients")+
  scale_y_discrete(labels = scales::wrap_format(14))+
  theme_bw()+
  theme(axis.text = element_text(color="black"))

<<<<<<< HEAD
p1 + p2
ggsave(here("figures", "coefplot.png"),
       device="png", height=4, 
       width=6, units="in", dpi=500)
=======
png("figures/coefplot.png", units="in", width=6, height=4, res=500)
p1 + p2
dev.off()
>>>>>>> bf4084247c05632565b8e98746a0fd6754471bc0

```
