---
title: "Replication Norris-Inglehart, ESS 1-7"
author: "Armin Schäfer"
date: "`r format(Sys.time(),  '%B %d, %Y')`"
output: html_document
---

In this file, I reproduce some of the analyses in Norrs/Inglehart (2019) with the European Social Survey rounds 1 to 7.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

```{r}
# Loading R-libraries.

if (!require("pacman")) install.packages("pacman")
pacman::p_load(essurvey, readr, tidyverse, Hmisc, dplyr, psych, naniar, here, scales,
               srvyr, fixest, corpcor, lme4, RCurl, ggpubr)

# useful function
tabl <- function(...) table(..., useNA="ifany")

# load the dataset created with "arb_ess_1-7.RMD"

ess_1_7 <- readRDS(here("processed-data", "arb_ess_1_7.RDS"))

```


Median age per cohort
```{r}

ess_1_7 %>%
  group_by(cohort)%>%
  summarise(medage = median(age, na.rm=T))

```

Cultural attitudes

Example: Gays and lesbians should be free to life their lives at the wish
```{r, message = FALSE, warning = FALSE}

p1 <- ess_1_7 %>%
  filter_at(vars(cohort, freehms),all_vars(!is.na(.)))%>%
  mutate(freehmsz = wt.scale(freehms, weight)) %>%
  group_by(cohort) %>%
  summarise(mfree = weighted.mean(freehmsz, weight, na.rm=T)) %>%
ggplot()+
  geom_hline(yintercept=0, color="grey70")+
  geom_point(aes(x = cohort, y = mfree),
             size=3)+
  labs(x = NULL, y = "Z-transformed scale\n")+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(-0.5,0.5)+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=10))


p2 <- ess_1_7 %>%
  filter_at(vars(cohort, freehms),all_vars(!is.na(.)))%>%
  group_by(cohort) %>%
  summarise(mfree = weighted.mean(freehms, weight, na.rm=T)) %>%
ggplot()+
  geom_point(aes(x = cohort, 
                 y = mfree),
             size=3)+
  geom_hline(yintercept=3, color="grey70")+
  ylim(1,5)+
  labs(x = NULL, y = "Original scale\n")+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
        labels=c("Interwar", "Boomers", 
                 "Gen X", "Millennials"))+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=10))

p3 <- ggarrange(p1, p2)

ggsave(here("figures", "freehms.png"),
       device="png", height=4, 
       width=7, units="in", dpi=500)


```


Attitudes in comparison: different scales
```{r}

p2_1 <- ess_1_7 %>%
  filter_at(vars(cohort, imueclt, imwbcnt, euftf),all_vars(!is.na(.)))%>%
  select(cohort, imueclt, imwbcnt, euftf, weight)%>%
  mutate(imuecltz = wt.scale(imueclt, weight),
         euftfz = wt.scale(euftf, weight),
         imwbcntz = wt.scale(imwbcnt, weight)) %>%
  group_by(cohort) %>%
  summarise("Mig. culture" = weighted.mean(imuecltz, weight, na.rm=TRUE),
            "EU integration" = weighted.mean(euftfz, weight, na.rm=TRUE),
            "Mig. economy" = weighted.mean(imwbcntz, weight, na.rm=TRUE)) %>%
  pivot_longer(-cohort, names_to = "items", values_to = "index") %>%
ggplot()+
  geom_hline(yintercept = 0, color="grey70")+
  geom_point(aes(x = cohort, y = index), size=2)+
  labs(y = "z-transformed scores\n",
       x = NULL,
       color = NULL)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(-0.3,0.3)+
  theme_bw()+
  theme(axis.text = element_text(color = "black", size=8),
        legend.position = "top")+
  facet_wrap(~items)


p2_2 <- ess_1_7 %>%
  filter_at(vars(cohort, imueclt, imwbcnt, euftf),all_vars(!is.na(.)))%>%
  select(cohort, imueclt, imwbcnt, euftf)%>%
  group_by(cohort) %>%
  summarise("Mig. culture" = weighted.mean(imueclt, na.rm=TRUE),
            "EU integration" = weighted.mean(euftf, na.rm=TRUE),
            "Mig. economy" = weighted.mean(imwbcnt, na.rm=TRUE)) %>%
  pivot_longer(-cohort, names_to = "items", values_to = "index") %>%
ggplot()+
  geom_hline(yintercept=5, color="grey70")+
  geom_point(aes(x = cohort, y = index), size=2)+
  labs(y = "Original scale\n",
       x = NULL,
       color = NULL)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(0,10)+
  theme_bw()+
  theme(axis.text = element_text(color = "black", size=8),
        legend.position = "top")+
  facet_wrap(~items)

p2_3 <- ggarrange(p2_1, p2_2, nrow=2)

ggsave(here("figures", "cultitems.png"),
       device="png", height=5, 
       width=7, units="in", dpi=500)


```


Attitudes in comparison: different scales
```{r}

p3_1 <- ess_1_7 %>%
  filter_at(vars(cohort, freehms, wmcpwrk),all_vars(!is.na(.)))%>%
  select(cohort, freehms, wmcpwrk, weight)%>%
  mutate(freehmsz = wt.scale(freehms, weight),
         wmcpwrkz = wt.scale(wmcpwrk, weight)) %>%
  group_by(cohort) %>%
  summarise(gay_rights = weighted.mean(freehmsz, weight, na.rm=TRUE),
            women_work = weighted.mean(wmcpwrkz, weight, na.rm=TRUE)) %>%
  pivot_longer(-cohort, names_to = "items", values_to = "index") %>%
ggplot()+
  geom_hline(yintercept = 0, color="grey70")+
  geom_point(aes(x = cohort, y = index), size=2)+
  labs(y = "z scores\n",
       x = NULL,
       color = NULL)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(-0.3,0.3)+
  theme_minimal()+
  theme(axis.text = element_text(color = "black", size=8),
        legend.position = "top")+
  facet_wrap(~items)


p3_2 <- ess_1_7 %>%
  filter_at(vars(cohort, freehms, wmcpwrk),all_vars(!is.na(.)))%>%
  select(cohort, freehms, wmcpwrk,weight)%>%
  group_by(cohort) %>%
  summarise(gay_rights = weighted.mean(freehms, weight, na.rm=TRUE),
            women_work = weighted.mean(wmcpwrk, weight, na.rm=TRUE)) %>%
  pivot_longer(-cohort, names_to = "items", values_to = "index") %>%
ggplot()+
  geom_hline(yintercept=3, color="grey70")+
  geom_point(aes(x = cohort, y = index), size=2)+
  labs(y = "Original scale\n",
       x = NULL,
       color = NULL)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(1,5)+
  theme_minimal()+
  theme(axis.text = element_text(color = "black", size=8),
        legend.position = "top")+
  facet_wrap(~items)

p3_3 <- ggarrange(p3_1, p3_2, nrow=2)

```


Regression with and without country dummies
```{r}

ess_1_7 <- ess_1_7 %>%
  filter(RC1!="NA") %>%
  mutate(authz = wt.scale(RC1, weight))

m1 <- feols(authz~ cohort | essround, data = ess_1_7)
m2 <- feols(authz~ cohort | essround + cntry, data = ess_1_7)

etable(m1, m2)

```


Authoritarian and libertarian values - z-transformed factors
```{r}

p1_1 <- ess_1_7 %>%
  filter_at(vars(cohort, RC1, RC2),all_vars(!is.na(.)))%>%
  mutate(authz = wt.scale(RC1, weight),
         libz = wt.scale(RC2, weight)) %>%
  group_by(cohort) %>%
  summarise(authoritarian = weighted.mean(authz, weight, na.rm=T),
            libertarian  = weighted.mean(libz, weight, na.rm =T)) %>% 
  select(cohort, authoritarian, libertarian) %>%
  pivot_longer(-cohort, names_to = "scale", values_to = "values") %>%
ggplot()+
  geom_line(aes(x = cohort, y = values, 
                group=scale,
                linetype=scale), size=1)+
  geom_point(aes(x = cohort, y = values), size=2)+
  geom_hline(yintercept=0,
             color = "grey70")+
  labs(x = NULL, y = "z scores",
       group = NULL, linetype = NULL)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
        labels=c("Interwar", "Boomers", "Gen X", "Millennials"))+
  scale_linetype_manual(values=c("twodash", "dotted"))+
  ylim(-.5,.5)+
  theme_bw()+
  theme(axis.text = element_text(color="black",
                                 size = 10))

ggsave(here("figures", "schwartz.png"),
       device="png", height=4, 
       width=6, units="in", dpi=500)

```


Authoritarian and libertarian values - orginal scales
```{r}

p3_1 <- ess_1_7 %>%
  select(cohort, weight, behave, secure, safety, tradition, rules) %>%
  filter(cohort!="NA") %>%
  pivot_longer(-(1:2), names_to="item", values_to="index") %>%
  group_by(cohort, item) %>%
  summarise(mindex = weighted.mean(index, weight, na.rm=T)) %>% 
ggplot()+
  geom_hline(yintercept = 3.5, color = "grey70")+
  geom_point(aes(x = cohort, y = mindex))+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", 
                            "Gen X", "Millennials"))+
  ylim(1,6)+
  labs(x = NULL,
       y = "Authoritarian values")+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=8))+
  coord_flip()+
  facet_wrap(~item, ncol=5)


p3_2 <- ess_1_7 %>%
  filter(cohort!="NA") %>%
  select(cohort, weight, surprise, adventure, creative, free, listen) %>%
  pivot_longer(-(1:2), names_to="item", values_to="index") %>%
  group_by(cohort, item) %>%
  summarise(mindex = weighted.mean(index, weight, na.rm=TRUE)) %>% 
ggplot(aes(x = cohort, y = mindex))+
  geom_hline(yintercept = 3.5, color="grey70")+
  geom_point(aes(x = cohort, y = mindex))+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers", "Gen X", "Millennials"))+
  ylim(1,6)+
  labs(x= NULL,
       y = "Libertarian values")+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=8))+
  coord_flip()+
  facet_wrap(~item, ncol=5)

p3_3 <- ggarrange(p3_1, p3_2, nrow=2)

ggsave(here("figures", "schwartzitems.png"),
       device="png", height=5, 
       width=7, units="in", dpi=500)

```

Average authoritarian and liberal values: Averages and violin plot
```{r}
av1 <- ess_1_7 %>%
  filter(cohort!="NA") %>%
  group_by(cohort) %>%
  summarise(mauth = weighted.mean(schwartzauth, weight, na.rm=TRUE)) %>% 
ggplot()+
  geom_violin(data = subset(ess_1_7, cohort!="NA"),
              aes(x = cohort, y = schwartzauth), 
              color = "grey80",
              fill = "grey80",
              adjust = 1.3)+
  geom_point(aes(x = cohort, y = mauth), color = "black")+
  geom_hline(yintercept = 50, color = "grey70")+
scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", 
                            "Boomers", 
                            "Gen X", 
                            "Millennials"))+
  #scale_y_continuous(limits = c(1,6),
  #                   breaks = c(1,2,3,4,5,6))+
  labs(x = NULL, y = "Authoritarian values\n")+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=8))

av2 <- ess_1_7 %>%
  filter(cohort!="NA") %>%
  group_by(cohort) %>%
  summarise(mlib = mean(schwartzlib, na.rm=TRUE),
            ci = list(mean_cl_normal(schwartzlib) %>% 
                        rename(mean=y, lwr=ymin, upr=ymax))) %>% 
  unnest(ci) %>%
ggplot()+
  geom_violin(data = subset(ess_1_7, cohort!="NA"), 
              aes(x = cohort, y = schwartzlib), 
              color = "grey80",
              fill = "grey80")+
  geom_pointrange(aes(x = cohort, y = mlib,
                      ymin = lwr, ymax = upr), color = "red")+
  geom_hline(yintercept = 50, color = "grey70")+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", 
                            "Boomers", 
                            "Gen X", 
                            "Millennials"))+
  #scale_y_continuous(limits = c(1,6),
  #                  breaks = c(1,2,3,4,5,6))+
  labs(x = NULL, y = "Libertarian values")+
  theme_minimal()+
  theme(axis.text = element_text(color="black", size=8))

```

Attitudes by country
```{r}
cntryplot <- ess_1_7 %>%
  filter(cohort!="NA") %>%
  group_by(cohort, cntry) %>%
  summarise(mauth = weighted.mean(schwartzauth, weight, na.rm=TRUE),
            mlib = weighted.mean(schwartzlib, weight, na.rm = TRUE)) %>% 
ggplot()+
  geom_point(aes(x = cohort, y = mauth), color = "red")+
  geom_point(aes(x = cohort, y = mlib), color = "blue")+
  geom_hline(yintercept = 50, color = "grey70")+
scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", 
                            "Boomers", 
                            "Gen X", 
                            "Millennials"))+
  labs(x = NULL, y = "Authoritarian values\n")+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=8))+
  facet_wrap(~cntry)

```


Populist values (sum of trstplt, trstprl, trstprt scaled to 100)
```{r}

by_country <- ess_1_7 %>%
  group_by(cntry) %>%
  summarise(mpop = 100-weighted.mean(populist, weight, na.rm=TRUE))

by_cohort <- ess_1_7 %>%
  group_by(cohort) %>%
  summarise(mpop = 100-weighted.mean(populist, weight, na.rm=TRUE))

plotpop1 <- ess_1_7 %>%
  filter(cohort!="NA") %>%
  group_by(cohort, cntry) %>%
  summarise(mpop = 100-weighted.mean(populist, weight, na.rm=TRUE)) %>%
ggplot()+
  geom_jitter(aes(x = cohort, y = mpop,
                  color=cntry),
              color="grey50", 
             size=2,
             width=0.1,
             alpha = 0.5)+
  geom_point(data = by_cohort, aes(x=cohort, y=mpop),
             size=3.5,
             color="black")+
  labs(x = NULL, y = "Populist attitudes\n (reversed political trust scale)")+
  geom_hline(yintercept=50, color="grey70")+
  ylim(0,100)+
  scale_x_discrete(breaks=c("1", "2", "3", "4"),
                   labels=c("Interwar", "Boomers",
                            "Gen X", "Millennials"))+
  theme_bw()+
  theme(axis.text = element_text(color="black", size=10))

ggsave(here("figures", "plotpop_1.png"),
       device="png", height=5, 
       width=7, units="in", dpi=500)

```

Regression model populist attitudes
```{r}

forreg <- ess_1_7 %>%
  filter_at(vars(cohort, RC1, RC2, populist),all_vars(!is.na(.)))%>%
  mutate(authz = wt.scale(RC1, weight),
         libz = wt.scale(RC2, weight),
         popz = wt.scale(populist, weight)) 
interceptonly <- lmer(formula = populist ~ 1 + (1|cntry),
                           data    = ess_1_7)

summary(interceptonly)

icc <- 110.5/(110.5+415.8)
icc # Twenty percent of the variance are explained at the country level

m1 <- lmer(formula = populist~ 1 + cohort + gender + urban + relig + factor(class5) +
             schwartzauth + (1 |cntry), data = ess_1_7)

summary(m1)

m1 <- feols(populist~ cohort + schwartzauth | cntry, data = forreg)
etable(m1)

m2 <- lm(populist~ cohort + gender + urban + relig + factor(class5) +
             schwartzauth, data = ess_1_7)
summary(m2)

```


